### 2022-07-26 src/text/1-introduction.tex
Consistently use Gammapy shortcut

\begin{figure*}[t]
	\centering
	\includegraphics[height=\textwidth, angle=270]{static/gammapy-big-picture}
    \caption{ Gammapy is a Python package
		for high-level \gammaray data analysis. Using event lists, exposures and point
		spread functions as input you can use it to generate science results such as
		images, spectra, light curves or source catalogs. So far it has been used to
		simulate and analyse \hess, \cta,  and \fermi data. As demonstrated
		by several studies, it will also be applied to e.g., \veritas, \magic or \hawc data
		in the future. }
	\label{fig:big-picture}

@@

%% Gammapy: concept and goals
%%% Question: starting date? Is is relevant, what is the real start?
The Gammapy project started in 2015 with the objective of building a flexible and
efficient software library for very high energy \gammaray data analysis \citep{gammapy_2015}.
This is possible thanks to the definition of a common data format such as \gadf, and thanks to
an open development with a broad community of contributors.

%Gammapy is built on the scientific Python stack and makes use of Astropy.

%% Gammapy: from first developments to validation and selection
The H.E.S.S. collaboration released a limited test dataset (about 50 hours of
observations taken between 2004 and 2008) based  on the \gadf DL3 format \citep{HESS_DR1}.
This data release served as a basis for analysis tools validation \cite[see e.g.]{Mohrmann2019}.

TODO: Figure 1: Data -> Gammapy -> Spectra etc with some details

%% Paper outline
In this article, we describe the general structure of the Gammapy package and
its main concepts. In Section \ref{sec:analysis-workflow-overview}, we present
the data analysis workflow in very high energy \gammaray astronomy. Then we
show how this workflow is reflected in the structure of the Gammapy package and
describe the various subpackages it contains. Section \ref{sec:applications}
presents a number of applications, while Section \ref{sec:gammapy-project}
discusses the project organization.

>>>>

\begin{figure*}[t]
	\centering
	\includegraphics[height=\textwidth, angle=270]{static/gammapy-big-picture}
    \caption{ \gammapy is a Python package
		for high-level \gammaray data analysis. Using event lists, exposures and point
		spread functions as input you can use it to generate science results such as
		images, spectra, light curves or source catalogs. So far it has been used to
		simulate and analyse \hess, \cta,  and \fermi data. As demonstrated
		by several studies, it will also be applied to e.g., \veritas, \magic or \hawc data
		in the future. }
	\label{fig:big-picture}

@@

%% \gammapy: concept and goals
%%% Question: starting date? Is is relevant, what is the real start?
The \gammapy project started in 2015 with the objective of building a flexible and
efficient software library for very high energy \gammaray data analysis \citep{gammapy_2015}.
This is possible thanks to the definition of a common data format such as \gadf, and thanks to
an open development with a broad community of contributors.

%\gammapy is built on the scientific Python stack and makes use of Astropy.

%% \gammapy: from first developments to validation and selection
The \hess collaboration released a limited test dataset (about 50 hours of
observations taken between 2004 and 2008) based  on the \gadf DL3 format \citep{HESS_DR1}.
This data release served as a basis for analysis tools validation \cite[see e.g.]{Mohrmann2019}.

TODO: Figure 1: Data -> \gammapy -> Spectra etc with some details

%% Paper outline
In this article, we describe the general structure of the \gammapy package and
its main concepts. In Section \ref{sec:analysis-workflow-overview}, we present
the data analysis workflow in very high energy \gammaray astronomy. Then we
show how this workflow is reflected in the structure of the \gammapy package and
describe the various subpackages it contains. Section \ref{sec:applications}
presents a number of applications, while Section \ref{sec:gammapy-project}
discusses the project organization.

### 2022-07-26 src/text/2-package-subsections/estimators.tex
Use shortcuts more consistently
Consistently use Gammapy shortcut

\begin{table*}
    \begin{center}
        \begin{tabular}{lll}
         \hline
         Type & Description & Unit Equivalency \\
         \hline
         dnde & Differential flux at a given energy & $\mathrm{TeV^{-1}~cm^{2}~s^{-1}}$ \\
         e2dnde & Differential flux at a given energy  & $\mathrm{TeV~cm^{2}~s^{-1}}$ \\
         flux & Integrated flux in agiven energy range & $\mathrm{cm^{2}~s^{-1}}$ \\
         eflux & Integrated energy flux in agiven energy range & $\mathrm{erg~cm^{2}~s^{-1}}$\\
         \hline
        \end{tabular}
    \end{center}
    \label{tab:sed_types}
    \caption{Definition of sed types.}
\end{table*}

Both result objects support the possibility to serialise
the data into multiple formats. This includes the
GADF SED (reference?) format, FITS based ND sky-maps
and formats compatible with Astropy's \code{Table} and
\code{BinnedTimeSeries} data structures. This allows
users to directly further analyse the results, e.g.
standard algorithms for time analysis such as
computing Lomb-Scargle periodograms or Bayesian
blocks for time series. So far \gammapy does not
support "unfolding" of \gammaray spectra.
Methods for this will be implemented in future version of \gammapy.

>>>>

\begin{table*}
    \begin{center}
        \begin{tabular}{lll}
         \hline
         Type & Description & Unit Equivalency \\
         \hline
         dnde & Differential flux at a given energy & $\mathrm{TeV^{-1}~cm^{2}~s^{-1}}$ \\
         e2dnde & Differential flux at a given energy  & $\mathrm{TeV~cm^{2}~s^{-1}}$ \\
         flux & Integrated flux in agiven energy range & $\mathrm{cm^{2}~s^{-1}}$ \\
         eflux & Integrated energy flux in agiven energy range & $\mathrm{erg~cm^{2}~s^{-1}}$\\
         \hline
        \end{tabular}
    \end{center}
    \label{tab:sed_types}
    \caption{Definition of the different SED types supported in \gammapy.}
\end{table*}

Both result objects support the possibility to serialise
the data into multiple formats. This includes the
\gadf SED (reference?) format, FITS based ND sky-maps
and formats compatible with Astropy's \code{Table} and
\code{BinnedTimeSeries} data structures. This allows
users to directly further analyse the results, e.g.
standard algorithms for time analysis such as
computing Lomb-Scargle periodograms or Bayesian
blocks for time series. So far \gammapy does not
support "unfolding" of \gammaray spectra.
Methods for this will be implemented in future version of \gammapy.

### 2022-07-26 src/text/2-package.tex
Use title case consistently
Consistently use Gammapy shortcut

\section{\gammaray Data Analysis}
\label{sec:gammaray-data-analysis}

@@

The counts data and the reduced IRFs in the form of maps are bundled into dataset objects
that represent the data level 4 (DL4). They can be written to
disk, in a format specific to Gammapy to allow users to read them back at any time later
for modeling and fitting.

@@

The next step is then typically to model and fit the datasets, either
individually, or in a joint likelihood analysis. For this purpose Gammapy
provides a uniform interface to multiple fitting backends. It also provides
a variety of :ref:`built in models <model-gallery>`. This includes spectral,
spatial and temporal model classes to describe the \gammaray emission in the sky.
Where spectral models can be simple analytical models or more complex ones from radiation
mechanisms of accelerated particle populations (e.g. inverse Compton or $\pi^{o}$ decay).
Independently or subsequently to the global modelling, the data can be
re-grouped to compute flux points, light curves and flux as well as significance
maps in energy bands.

\section{Gammapy Package}
\label{sec:gammapy-package}
\subsection{Overview}
\label{ssec:overview}
%
\begin{figure*}[ht!]
	\centering
	\includegraphics[width=1.\textwidth]{figures/data_flow.pdf}
	\caption{
		Gammapy sub-package structure and data analysis workflow. The top row
        defines the groups for the different data levels and reduction steps
        from raw gamma-like events on the left, to high level science products
        on the right. The direction of the data flow is illustrated with the
        grey arrows. The gray folder icons represent the different sub-packages
        in \gammapy and their names. Below each icon there is a list of the most
        important objects defined in the sub-package.
    }
	\label{fig:data_flow}
\end{figure*}
%
The \gammapy package is structured into multiple sub-packages. The definition
of the content of the different sub-packages follows mostly the stages in the
data reduction workflow described in the previous section. Sub-packages
either contain data structures representing data at different data reduction
levels or contain algorithms to transition between the different data reduction
levels.

>>>>

\section{Gamma-ray Data Analysis}
\label{sec:gammaray-data-analysis}

@@

The counts data and the reduced IRFs in the form of maps are bundled into dataset objects
that represent the data level 4 (DL4). They can be written to
disk, in a format specific to \gammapy to allow users to read them back at any time later
for modeling and fitting.

@@

The next step is then typically to model and fit the datasets, either
individually, or in a joint likelihood analysis. For this purpose \gammapy
provides a uniform interface to multiple fitting backends. It also provides
a variety of :ref:`built in models <model-gallery>`. This includes spectral,
spatial and temporal model classes to describe the \gammaray emission in the sky.
Where spectral models can be simple analytical models or more complex ones from radiation
mechanisms of accelerated particle populations (e.g. inverse Compton or $\pi^{o}$ decay).
Independently or subsequently to the global modelling, the data can be
re-grouped to compute flux points, light curves and flux as well as significance
maps in energy bands.

\section{\gammapy Package}
\label{sec:gammapy-package}
\subsection{Overview}
\label{ssec:overview}
%
\begin{figure*}[ht!]
	\centering
	\includegraphics[width=1.\textwidth]{figures/data_flow.pdf}
	\caption{
		\gammapy sub-package structure and data analysis workflow. The top row
        defines the groups for the different data levels and reduction steps
        from raw gamma-like events on the left, to high level science products
        on the right. The direction of the data flow is illustrated with the
        grey arrows. The gray folder icons represent the different sub-packages
        in \gammapy and their names. Below each icon there is a list of the most
        important objects defined in the sub-package.
    }
	\label{fig:data_flow}
\end{figure*}
%
The \gammapy package is structured into multiple sub-packages. The definition
of the content of the different sub-packages follows mostly the stages in the
data reduction workflow described in the previous section. Sub-packages
either contain data structures representing data at different data reduction
levels or contain algorithms to transition between the different data reduction
levels.

### 2022-07-26 src/text/4-project.tex
Use title case consistently
Consistently use Gammapy shortcut

\section{Gammapy project} \label{sec:gammapy-project}

In this section, we provide an overview of the organization of the Gammapy
project. We briefly describe the main roles and responsibilities within the
team, as well as the technical infrastructure designed to facilitate the
development and maintenance of Gammapy as a high quality software. We use
common tools and services for software development of Python open-source
projects, code review, testing, package distribution and user support, with a
customized solution for a versioned thoroughly tested documentation in the form
of user-friendly playable tutorials. This section concludes with an outlook on
the roadmap for future directions.

\subsection{Organizational structure}
\label{ssec:organizational-structure}

Gammapy is a well-organised international open-source project with a broad
developer base and the presence of commitments from strong groups and
institutes in the very high-energies astrophysics
domain\footnote{\url{https://gammapy.org/team.html}}. The main development
roadmaps are discussed and validated by a Coordination Committee, composed of
representatives of the main contributing laboratories. This committee is
chaired by a Project Manager and his deputy while two Lead Developers manage
the development strategy and organise technical activities. Because of this
institutionally driven characteristic, the permanent staff and commitment of
supporting institutes ensure the continuity of the executive teams. A core team
of developers from the contributing laboratories is in charge of the regular
development, which benefits from punctual contributions of the community at
large.

\subsection{Technical infrastructure}
\label{ssec:technical-infrastructure}

Gammapy follows an open-source and open-contribution development model based on
the cloud repository service \github. A \github organization
\textit{gammapy}\footnote{\url{https://\github.com/gammapy}} hosts different
repositories related with the project. The software codebase may be found in
the \textit{gammapy} repository (see Table~\ref{table:codestats:data} and
Figure~\ref{fig:codestats:lang} for code lines statistics). We make extensive
use of the pull request system to discuss and review code contributions.

@@

\subsection{Software distribution}
\label{ssec:software-distribution}

Gammapy is distributed for Linux, Windows and Mac environments, and installed
in the usual way for Python packages. Each stable release is uploaded to the
Python package index\footnote{\url{https://pypi.org}} and as a binary package
to the \textit{conda-forge} and \textit{astropy} Anaconda
repository\footnote{\url{https://anaconda.org/anaconda/repo}} channels. At this
time, Gammapy is also available as a Debian Linux
package\footnote{\url{https://packages.debian.org/sid/python3-gammapy}}. We
recommend installing the software using the \textit{conda} installation process
with an environment definition file that we provide, so to work within a
virtual isolated environment with additional useful packages and ensure
reproducibility.

\gammapy is indexed in Astronomy Source Code
Library\footnote{\url{https://ascl.net/1711.014}} and
Zenodo\footnote{\url{https://zenodo.org/record/5721467o}} digital libraries for
software. The Zenodo record is synchronised with the codebase \github repository
so that every release triggers the update of the versioned record. In addition,
the next release of Gammapy will be added to the Open-source scientific
Software and Service Repository\footnote{\url{https://projectescape.eu/ossr}}
and indexed in the European Open Science Cloud
catalog\footnote{\url{https://eosc-portal.eu}}.

\subsection{Documentation and user-support}
\label{ssec:documentation-and-user-support}
Gammapy provides its user community with a tested and versioned up-to-date
on-line
documentation\footnote{\url{https://docs.gammapy.org}}~\citep{2019ASPC..523..357B}
built with Sphinx\footnote{\url{https://www.sphinx-doc.org}} scanning the
codebase Python scripts, as well as a set of RST files and Jupyter notebooks.
The documentation includes a handwritten user guide, a set of executable
tutorials, and a reference to the API automatically extracted from the code and
docstrings. The Gammapy code snippets present in the documentation are tested
in different environments using our continuous integration (CI) workflow based
on \github actions.

@@

We have also set up a solution for users to share recipes as Jupyter notebooks
that do not fit in the Gammapy core documentation but which may be relevant as
specific use cases. Contributions happen via pull requests to the
\textit{gammapy-recipes} \github repository and merged after a short review. All
notebooks in the repository are tested and published in the Gammapy recipes
webpage\footnote{\url{https://gammapy.\github.io/gammapy-recipes}} automatically
using \github actions.

A growing community of users is gathering around the Slack
messaging\footnote{\url{https://gammapy.slack.com}} and \github
discussions\footnote{\url{https://\github.com/gammapy/gammapy/discussions}}
support forums, providing valuable feedback on the Gammapy functionalities,
interface and documentation. Other communication channels have been set like
mailing lists, a Twitter account\footnote{\url{https://twitter.com/gammapyST}},
regular public coding sprint meetings, hands-on session within collaborations,
weekly development meetings, etc.

\subsection{Proposals for improving \gammapy}
\label{ssec:pigs}
An important part of \gammapy's development organisation is the support
for "Proposals for improving Gammapy" (PIG). This system is very much
inspired by Python's PEP\footnote{\url{https://peps.python.org/pep-0001/}}
and Astropy's APE \citep{greenfield_perry_2013} system.
PIG are self-contained documents which outline a set of larger
changes to the \gammapy code base. This includes larger feature additions,
code and package restructuring and maintenance as well as changes related
to the organisational structure of the \gammapy project. PIGs can be proposed
by any person in or outside the project and by multiple authors. They
are presented to the \gammapy developer community in a pull request
on \github and the undergo a review phase in which changes and
improvements to the document are proposed and implemented. Once the PIG
document is in a final state it is presented to the \gammapy
coordination committee, which takes the final decision on the
acceptance or rejection of the proposal. Once accepted, the proposed
change are implemented by \gammapy developers in a series of
individual contributions via pull requests. A list of all proposed
PIG documents is available in the \gammapy online documentation
\footnote{\url{https://docs.gammapy.org/0.20/development/pigs/index.html}}.

>>>>

\section{\gammapy Project} \label{sec:gammapy-project}

In this section, we provide an overview of the organization of the \gammapy
project. We briefly describe the main roles and responsibilities within the
team, as well as the technical infrastructure designed to facilitate the
development and maintenance of \gammapy as a high quality software. We use
common tools and services for software development of Python open-source
projects, code review, testing, package distribution and user support, with a
customized solution for a versioned thoroughly tested documentation in the form
of user-friendly playable tutorials. This section concludes with an outlook on
the roadmap for future directions.

\subsection{Organizational Structure}
\label{ssec:organizational-structure}

\gammapy is a well-organised international open-source project with a broad
developer base and the presence of commitments from strong groups and
institutes in the very high-energies astrophysics
domain\footnote{\url{https://gammapy.org/team.html}}. The main development
roadmaps are discussed and validated by a Coordination Committee, composed of
representatives of the main contributing laboratories. This committee is
chaired by a Project Manager and his deputy while two Lead Developers manage
the development strategy and organise technical activities. Because of this
institutionally driven characteristic, the permanent staff and commitment of
supporting institutes ensure the continuity of the executive teams. A core team
of developers from the contributing laboratories is in charge of the regular
development, which benefits from punctual contributions of the community at
large.

\subsection{Technical Infrastructure}
\label{ssec:technical-infrastructure}

\gammapy follows an open-source and open-contribution development model based on
the cloud repository service \github. A \github organization
\textit{gammapy}\footnote{\url{https://\github.com/gammapy}} hosts different
repositories related with the project. The software codebase may be found in
the \textit{gammapy} repository (see Table~\ref{table:codestats:data} and
Figure~\ref{fig:codestats:lang} for code lines statistics). We make extensive
use of the pull request system to discuss and review code contributions.

@@

\subsection{Software Distribution}
\label{ssec:software-distribution}

\gammapy is distributed for Linux, Windows and Mac environments, and installed
in the usual way for Python packages. Each stable release is uploaded to the
Python package index\footnote{\url{https://pypi.org}} and as a binary package
to the \textit{conda-forge} and \textit{astropy} Anaconda
repository\footnote{\url{https://anaconda.org/anaconda/repo}} channels. At this
time, \gammapy is also available as a Debian Linux
package\footnote{\url{https://packages.debian.org/sid/python3-gammapy}}. We
recommend installing the software using the \textit{conda} installation process
with an environment definition file that we provide, so to work within a
virtual isolated environment with additional useful packages and ensure
reproducibility.

\gammapy is indexed in Astronomy Source Code
Library\footnote{\url{https://ascl.net/1711.014}} and
Zenodo\footnote{\url{https://zenodo.org/record/5721467o}} digital libraries for
software. The Zenodo record is synchronised with the codebase \github repository
so that every release triggers the update of the versioned record. In addition,
the next release of \gammapy will be added to the Open-source scientific
Software and Service Repository\footnote{\url{https://projectescape.eu/ossr}}
and indexed in the European Open Science Cloud
catalog\footnote{\url{https://eosc-portal.eu}}.

\subsection{Documentation and User-support}
\label{ssec:documentation-and-user-support}
\gammapy provides its user community with a tested and versioned up-to-date
on-line
documentation\footnote{\url{https://docs.gammapy.org}}~\citep{2019ASPC..523..357B}
built with Sphinx\footnote{\url{https://www.sphinx-doc.org}} scanning the
codebase Python scripts, as well as a set of RST files and Jupyter notebooks.
The documentation includes a handwritten user guide, a set of executable
tutorials, and a reference to the API automatically extracted from the code and
docstrings. The \gammapy code snippets present in the documentation are tested
in different environments using our continuous integration (CI) workflow based
on \github actions.

@@

We have also set up a solution for users to share recipes as Jupyter notebooks
that do not fit in the \gammapy core documentation but which may be relevant as
specific use cases. Contributions happen via pull requests to the
\textit{gammapy-recipes} \github repository and merged after a short review. All
notebooks in the repository are tested and published in the \gammapy recipes
webpage\footnote{\url{https://gammapy.\github.io/gammapy-recipes}} automatically
using \github actions.

A growing community of users is gathering around the Slack
messaging\footnote{\url{https://gammapy.slack.com}} and \github
discussions\footnote{\url{https://\github.com/gammapy/gammapy/discussions}}
support forums, providing valuable feedback on the \gammapy functionalities,
interface and documentation. Other communication channels have been set like
mailing lists, a Twitter account\footnote{\url{https://twitter.com/gammapyST}},
regular public coding sprint meetings, hands-on session within collaborations,
weekly development meetings, etc.

\subsection{Proposals for Improving \gammapy}
\label{ssec:pigs}
An important part of \gammapy's development organisation is the support
for "Proposals for improving \gammapy" (PIG). This system is very much
inspired by Python's PEP\footnote{\url{https://peps.python.org/pep-0001/}}
and Astropy's APE \citep{greenfield_perry_2013} system.
PIG are self-contained documents which outline a set of larger
changes to the \gammapy code base. This includes larger feature additions,
code and package restructuring and maintenance as well as changes related
to the organisational structure of the \gammapy project. PIGs can be proposed
by any person in or outside the project and by multiple authors. They
are presented to the \gammapy developer community in a pull request
on \github and the undergo a review phase in which changes and
improvements to the document are proposed and implemented. Once the PIG
document is in a final state it is presented to the \gammapy
coordination committee, which takes the final decision on the
acceptance or rejection of the proposal. Once accepted, the proposed
change are implemented by \gammapy developers in a series of
individual contributions via pull requests. A list of all proposed
PIG documents is available in the \gammapy online documentation
\footnote{\url{https://docs.gammapy.org/0.20/development/pigs/index.html}}.

### 2022-07-26 src/text/6-acknowledgements.tex
Reformat text files
Introduce GtiHub shortcut
Consistently use Gammapy shortcut

	We would like to thank the \texttt{Numpy}, \texttt{Scipy}, \texttt{IPython} and
	\texttt{Matplotlib} communities for providing their packages which are
	invaluable to the development of Astropy. We thank the GitHub team for
	providing us with an excellent free development platform. We also are grateful
	to Read the Docs (\ReadthedocsUrl), and Travis (\TravisUrl) for providing free
	documentation hosting and testing respectively. Finally, we would like to thank
	all the Gammapy users that have provided feedback and submitted bug reports.

>>>>

	We would like to thank the \texttt{Numpy}, \texttt{Scipy}, \texttt{IPython} and
	\texttt{Matplotlib} communities for providing their packages which are
	invaluable to the development of Astropy. We thank the \github team for
	providing us with an excellent free development platform. We also are grateful
	to Read the Docs (\ReadthedocsUrl), and Travis (\TravisUrl) for providing free
	documentation hosting and testing respectively. Finally, we would like to thank
	all the \gammapy users that have provided feedback and submitted bug reports.
